### 2021-02-19 architecture.tex
Improve wording in Architecture

The architecture is designed so that the recommender can be trained
on the results of \gls{atp} executions on various problems with random precedences.
The recommender is parameterized by weight tensors
whose values can be trained by gradient descent.
\todo{Ensure we have defined ``gradient descent''.}
\todo{MS: chapu ucel (a vidim, ze dalsi veta navazuje na ``parameters'', ale je to kostrbate.
Recommender = system, chapu jako program. Programy vetsinou neparametrizujeme tenzorama a gradient descent zije na jine urovni,
nez aby parametrizovane programy primo ucil. Uceni je spis nejaka technologie, kterou recommender pouziva, aby ... ``produces precedences
that are likely to yield a successful proof''.  }
The goal of the training is to find parameter values
such that the recommender produces precedences
that are likely to yield a successful proof in few iterations of the saturation loop.
\todo{FB: Too vague? MS: v tomhle momente mi to prijde dobry!}

The recommender consists of modules that perform specific subtasks (see \cref{fig:architecture}):
\begin{enumerate}
\item Graphifier\todo{MS: Zatim nevidim Graphifier ve \cref{fig:architecture}.} converts the input \gls{cnf} problem into a problem graph.
\item \gls{gcn} converts a problem graph into symbol embeddings.
The \gls{gcn} contains one or more convolution layers.
\item Projection converts the symbol embeddings into a vector of symbol costs.
\item Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{enumerate}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... }

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous directed graph}.\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
\todo{FB: Consider removing the footnote because it is basically useless.}
\todo{MS: Footnote asi remove, ale neexistuje citace pro ``heterogeneous directed graph''?}
Each of the nodes is labeled with a node type,
and each edge is labeled with an edge type,
informing\todo{MS: maybe better ``defining''?} the heterogeneous nature of the graph.
The nodes of the graph correspond to various elements that constitute the \gls{cnf} formula,
such as predicate symbols, atoms, and clauses.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the incidence relation between a clause and one of its (literal's) atoms,
or the relation between an atom and its predicate symbol.
\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.

@@

The graph representation has the following properties:
\begin{itemize}
\item For each relation $r$, its inverse $\inv{r}$ is also present in the graph,
represented by a different edge type (unless $r = \inv{r}$ is an identity).
% Note: The relation is considered an identity if it is a subset of the complete identity relation.
\todo{MS: tohle jsem nepochopil. Nemuzeme rict, ze (az na nejakou vyjimkou) jsou edges v obou smerech?}
\item Lossless: The original problem can be faithfully reconstructed from the corresponding graph representation
(up to logical equivalence).
\item Signature agnostic: Renaming the symbols and variables in the input problem yields an isomorphic graph.
\item A singleton node of type \ntype{problem} is connected to all the clauses of the problem.
\item The polarity of the literals is expressed by the type of the edge (\epos{} or \eneg{})
connecting the respective atom to the clause it occurs in.
\item For every non-equality atom and term, the order of its arguments is captured by a sequence of \ntype{argument} nodes chained by edges \cite{Rawson2020}.
\item The two operands of equality are not ordered.
This reflects the symmetry of equality.
\item Term-sharing is employed:\todo{MS: Mozna presneji ``perfect sub-expression sharing''? 
Predpokladam ze dva ruzne ground literaly p(c) tam budou taky jen jendnou?}
Redundant atoms and terms share a node representation \cite{}.
Note that since each variable is bound by a clause,
ground terms may be shared across clauses,
but non-ground terms may only be shared within a clause.
\todo{FB: Remove as of little interest?}
\end{itemize}
\todo{FB: Reference appendix if we describe the representation in more detail there.}
\todo[inline]{MS: Filosoficka k ``can be represented''; mozna by se mohlo nekde objevit:
Snazime se do struktury HIN ``otisknout'' celou strukturu CNF (jak jsem rikal,
znamena to, ze zobrazeni je proste), abychom siti nevzali moznost ``vedet o problemu
uplne vsechno'' (az na veci, na kterych nezalezi jako poradi literaly v klauzulich, etc).
Ta filosoficka poznamka je o tom, ze pro dobre uceni tohle vlastne vubec nemusi byt treba,
ze nejaky ``hrubsi otisk'' by mohl stacit tez (a byt treba mensi), ale my nevime, jak takovy zvolit. }

@@

Note here that the symbol precedence heuristics commonly used in practice
conform to this scheme:
they order the symbols by a numeric property, such as arity or number of occurrences.

@@

The symbol cost model, given a \gls{cnf} problem $P = (\Sigma, \mathit{Cl})$,
assigns each symbol in $s \in \Sigma$ a real value $c(s)$.
It first uses a symbol embedding model to assign each symbol an embedding.
The resulting symbol cost is a dot product of the symbol embedding with a trainable weight vector:

>>>>

The architecture is designed so that the recommender can be trained
on the results of \gls{atp} executions on various problems with random precedences.
Since the recommender contains a neural network,
it is parameterized by weight tensors
whose values can be trained by gradient descent.
\todo{Ensure we have defined ``gradient descent''.}
\todo{MS: chapu ucel (a vidim, ze dalsi veta navazuje na ``parameters'', ale je to kostrbate.
Recommender = system, chapu jako program. Programy vetsinou neparametrizujeme tenzorama a gradient descent zije na jine urovni,
nez aby parametrizovane programy primo ucil. Uceni je spis nejaka technologie, kterou recommender pouziva, aby ... ``produces precedences
that are likely to yield a successful proof''.  }
The goal of the training is to find parameter values
such that the recommender produces precedences
that are likely to yield a successful proof in few iterations of the saturation loop.
\todo{FB: Too vague? MS: v tomhle momente mi to prijde dobry!}

The recommender consists of modules that perform specific subtasks (see \cref{fig:architecture}):
\begin{enumerate}
\item Graphifier\todo{MS: Zatim nevidim Graphifier ve \cref{fig:architecture}.} converts the input \gls{cnf} problem into a problem graph.
\item \gls{gcn} converts a problem graph into symbol embeddings.
%The \gls{gcn} contains one or more convolution layers.
\item Projection converts the symbol embeddings into a vector of symbol costs.
\item Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{enumerate}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... }

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous directed graph}.\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
\todo{FB: Consider removing the footnote because it is basically useless.}
\todo{MS: Footnote asi remove, ale neexistuje citace pro ``heterogeneous directed graph''?}
Each of the nodes is labeled with a node type,
and each edge is labeled with an edge type,
informing\todo{MS: maybe better ``defining''?} the heterogeneous nature of the graph.
Each node corresponds to one of the elements that constitute the \gls{cnf} formula,
such as a clause, an atom or a predicate symbol.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the incidence relation between a clause and one of its (literal's) atoms,
or the relation between an atom and its predicate symbol.
\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.

@@

The graph representation has the following properties:
\begin{itemize}
\item Lossless: The original problem can be faithfully reconstructed from the corresponding graph representation
(up to logical equivalence).
\item Signature agnostic: Renaming the symbols and variables in the input problem yields an isomorphic graph.
\item For each relation $r$, its inverse $\inv{r}$ is also present in the graph,
typically represented by a different edge type.
\todo{MS: tohle jsem nepochopil. Nemuzeme rict, ze (az na nejakou vyjimkou) jsou edges v obou smerech?}
\item A singleton node of type \ntype{problem} is connected to all the clauses of the problem.
\item The polarity of the literals is expressed by the type of the edge (\epos{} or \eneg{})
connecting the respective atom to the clause it occurs in.
\item For every non-equality atom and term, the order of its arguments is captured by a sequence of \ntype{argument} nodes chained by edges \cite{Rawson2020}.
\item The two operands of equality are not ordered.
This reflects the symmetry of equality.
\item Term-sharing is employed:\todo{MS: Mozna presneji ``perfect sub-expression sharing''? 
Predpokladam ze dva ruzne ground literaly p(c) tam budou taky jen jendnou?}
Redundant atoms and terms share a node representation \cite{}.
Note that since each variable is bound by a clause,
ground terms are shared across clauses,
but non-ground terms are only shared within the context of a clause.
\todo{FB: Remove as of little interest?}
\end{itemize}
\todo{FB: Reference appendix if we describe the representation in more detail there.}
\todo[inline]{MS: Filosoficka k ``can be represented''; mozna by se mohlo nekde objevit:
Snazime se do struktury HIN ``otisknout'' celou strukturu CNF (jak jsem rikal,
znamena to, ze zobrazeni je proste), abychom siti nevzali moznost ``vedet o problemu
uplne vsechno'' (az na veci, na kterych nezalezi jako poradi literaly v klauzulich, etc).
Ta filosoficka poznamka je o tom, ze pro dobre uceni tohle vlastne vubec nemusi byt treba,
ze nejaky ``hrubsi otisk'' by mohl stacit tez (a byt treba mensi), ale my nevime, jak takovy zvolit. }

@@

Note that the symbol precedence heuristics commonly used in practice
conform to this scheme:
they order the symbols by a numeric property, such as arity or number of occurrences.
\todo{MS: tohle se mi libi!}

@@

The symbol cost model, given a \gls{cnf} problem $P = (\Sigma, \mathit{Cl})$,
assigns each symbol in $s \in \Sigma$ a real value $c(s)$.
It first uses a symbol embedding model to assign each symbol $s$ an embedding $E(s)$.
The resulting symbol cost is a dot product of the symbol embedding with a trainable weight vector:
